%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper]{llncs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}
\usepackage{enumitem}
\usepackage{shuffle}
\usepackage{multirow}

\usepackage{tikz}
\usetikzlibrary{matrix, calc, arrows, shapes.callouts}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{complexity}
\input{graph}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \renewcommand{\floatpagefraction}{.9}
% \renewcommand{\textfraction}{.1}
\renewcommand{\arraystretch}{1.2}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathOperator{\LEFT}{\ell}
\DeclareMathOperator{\RIGHT}{r}
\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{s}}
\DeclareMathOperator{\SHUFFLE}{\bullet}
\DeclareMathOperator{\BINTOPERM}{\mathrm{btp}}
\DeclareMathOperator{\PERMTOBIN}{\mathrm{ptb}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{fact}{Fact}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\makeatletter
\newcommand{\pushright}[1]{\ifmeasuring@#1\else\omit\hfill$\displaystyle#1$\fi\ignorespaces}
\newcommand{\pushleft}[1]{\ifmeasuring@#1\else\omit$\displaystyle#1$\hfill\fi\ignorespaces}
\makeatother

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{%
Unshuffling Permutations}%

\author{%
  Samuele Giraudo \and
  St\'ephane Vialette
}% end author
\institute{%
  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC,
  F-77454, Marne-la-Vallée, France\\
  \email{samuele.giraudo@univ-mlv.fr} \\
  \email{vialette@univ-mlv.fr}
}% end institute
\date{\today}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
    A permutation is said to be a square if it is the union of two
    order-isomorphic patterns. The definition is intended to be the
    natural counterpart to the ordinary shuffle of words and languages.
    In this paper, we tackle the problem of recognizing square
    permutations from both the point of view of algebra and algorithmic.
    On the one hand, we present some algebraic properties of the shuffle
    product of permutations and exhibit a bijection between square
    $(213,231)$-avoiding permutations and square binary words. On the
    other hand, we prove that recognizing square permutations is \NPC.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%%
%% ---- Introduction ----
%%
\section{Introduction}
\label{section:Introduction}

Given three words $u$, $v$ and $w$, $u$ is said to a \emph{shuffle} of
$v$ and $w$ if it can be formed by interleaving the letters from $v$ and
$w$ in a way that maintains the left-to-right ordering of the letters
from each word (it is a classical textbook exercise to design an
efficient dynamic programming algorithm that determines whether $u$ is a
shuffle of $v$ and $w$). A word $u$ is said to be a \emph{square} (w.r.t.
the shuffle) if there exists a word $v$ such that $u$ is a shuffle of $v$
with itself.
\smallskip

This paper is intended to study a natural generalization of the shuffle
to permutations. Given permutations $\pi$, $\mu$ and $\nu$, $\pi$ is
said to be a shuffle of $\mu$ and $\nu$ if $\pi$ (viewed as a word) is
a shuffle of two words that are order-isomorphic to $\mu$ and $\nu$. For
example, $\pi$ is the shuffle of $\mu=$ and $\nu=$ as $\pi$. A
permutation $\pi$ is said to be a square (w.r.t. the shuffle) if there
exists a permutation $\mu$ such that $\pi$ is a shuffle of $\mu$ with
itself. For example, $\pi = 241356$ is a square as it is the shuffle of
$213$ with itself (\emph{e.g.},
$\pi = \dot{2}\ddot{4}\dot{1}\ddot{3}\ddot{5}\dot{6}$, where $\dot{i}$
and $\ddot{i}$), whereas $\pi = 4123$ is not.
The shuffle product defined and considered in this paper is different
from other existing and well-known shuffle products on permutations.
For instance, in~\cite{DHT:IJAC:2002}, the authors defined the
{\em shifted shuffle product}. In this product $\pi$ is a shuffle of
$\mu$ and $\nu$ if $\pi$ is the shuffle, as words, of $\mu$ and the
word obtained by incrementing all letters of $\nu$ by the size of $\mu$.
It is a simple exercise to prove that, give a permutation $\pi$, the
decision problem to test if $\pi$ is a square for the shifted shuffle
product is linear.
\smallskip

Formally, the shuffle of two words $u$ and $v$ of $A^*$ is the language
$u \shuffle v$ consisting of all words $u_1 v_1 u_2 v_2 \ldots u_k v_k$,
where $k \geq 0$ and the $u_i$ and the $v_i$ are the words  of $A^*$
such that $u = u_1 u_2 \ldots u_k$ and $v = v_1 v_2 \ldots v_k$. We
first briefly review the key results for words that arise in our context.
Given words $u$, $v_1$ and $v_2$, it can be tested in
$O\left(|u|^2 / \log(|u|)\right)$ time whether or not
$u \in v_1 \shuffle v_2$~\cite{Leeuwen:Nivat:IPL:1982}.
(To the best of our knowledge, the first $O(|u|^2)$ time algorithm a
ppeared in~\cite{Mansfield:DAM:1983}. This algorithm can easily be
extended to check in polynomial-time whether or not a word is the
shuffle of any fixed number of given words.) The shuffle $u \shuffle v$
of words $u$ and $v$ can be computed in
$O\left((|u|+|v|) \; \binom{|u|+|v|}{|u|}\right)$ time~\cite{Spehner:TCS:1986}.
An improvement and generalization has been proposed in
\cite{Allauzen:IGM:2000}, where it is proved that, given words
$u_1, u_2, \ldots, u_k$, the shuffle $\shuffle_{i=1}^{k} u_i$ can be
computed in
$O\left(\binom{|u_1|+|u_2|+\ldots+|u_k|}{|u_1|,|u_2|,\ldots,|u_k|}\right)$
time. Given words $u, v_1, v_2, \ldots, v_n \in A^*$, it is however \NPC
to decide whether or not $u \in \shuffle_{i=1}^{k} v_i$.
\cite{Mansfield:DAM:1983,Warmuth:Haussler:JCSS:1984}.
This remains true even if the alphabet has size $3$~\cite{Warmuth:Haussler:JCSS:1984}.
Of particular interest,
it is shown in~\cite{Warmuth:Haussler:JCSS:1984} that this problem remains
\NPC even if all
words $v_1, v_2, \ldots, v_n$ are identical,
thereby proving that,
for two words $u$ and $v$, it is \NPC to decide whether or not $u$
is in the iterated shuffle of $v$.
Again, this remains true even if the alphabet has size $3$.
As for squares, given a word $z$ it is \NPC to determine whether
there exists a word $x$ such that $z$ is the shuffle of $x$ with itself
\cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
\smallskip

$123$-avoiding (resp. $321$-avoiding) permutations are the union of two decreasing
(resp. increasing) patterns.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Notations ----
%%
\section{Notations}
\label{section:Notations}

For any nonnegative integer $n$, $[n]$ is the set $\{1, \dots, n\}$.
We follow the usual terminology on words~\cite{ChoffrutKarhumaki1997}.
Let us recall here the most important ones. The {\em empty word} is
denoted by $\epsilon$. If $u$ is a word, we denote by $\widetilde{u}$
the {\em mirror image} of $u$. If $P$ is a subset of $[|u|]$, $u_{|P}$
is the subword of $u$ consisting in the letters of $u$ at the positions
specified by the elements of $P$. The {\em shuffle} of two words $u$ and
$v$ is the set recursively defined by
$u \shuffle \epsilon = u = \epsilon \shuffle u$ and
$ua \shuffle vb = (u \shuffle vb)a \cup (ua \shuffle v)b$, were $a$ and
$b$ are letters. A word $u$ is a {\em square} if there exists a word $v$
such that $u$ belongs to $v \shuffle v$.

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. In this paper, permutations of a size $n$
are specified by words of length $n$ on the alphabet $[n]$ and without
multiple occurrence of a letter, so that all above definitions about
words remain valid on permutations. If $\pi$ is a permutation, we denote
by $\bar \pi$ (resp.~$\pi^{-1}$) the {\em complementary} (resp.
{\em inverse}) of $\pi$. If $u$ is a word of integers without multiple
occurrences of a same letter, $\STD(u)$ is the {\em standarized} of $u$,
that is the unique permutation of the same size as $u$ such that for all
$i, j \in [|u|]$, $u_i < u_j$ if and only if $\STD(u)_i < \STD(u)_j$. In
particular, the image of the map $\STD$ is the set $S$ of all
permutations. Two words $u$ and $v$ having the same standarized are
{\em order-isomorphic}. If $\sigma$ is a permutation, there is an
{\em occurrence} of (the {\em pattern}) $\sigma$ in $\pi$ if there is a
set $P$ of indexes of letters of $\pi$ such that $\sigma$ and $\pi_{|P}$
are order-isomorphic. When $\pi$ does not admit any occurrence of
$\sigma$, $\pi$ {\em avoids} $\sigma$.

For a graph $G$, we denote $\VS{G}$ as the set of vertices and
$\ES{G}$ as the set of edges.
Let $u = u_1 \, u_2 \,\ldots\, u_n$ be a word.
The \emph{clique graph associated to $u$}, denoted $K_u$,
is the complete graph on $\VS{K_u} = \{u_1, u_2, \ldots, u_n\}$.
The structure of this underlying graph is linear,
\emph{i.e.}, the set of vertices is equipped with a natural total order
$<$ defined by $u_i < u_j$ if and only if $i < j$.
(For a linear graph, we write $(u_i, u_j)$ instead of $\{u_i, u_j\}$
for an edge if it is clear from the context that $i < j$.)
Recall that two edges of a graph are \emph{independent} if they do not share
a common vertex, and that a \emph{matching} $\mathcal{M}$ in $G$ is a set of
pairwise independent edges.
A matching is \emph{perfect} if it covers all the vertices of the graph.
In case the set of vertices is equipped with a total order,
a matching $\mathcal{M}$ is said to be \emph{containment-free}
if there do not exist (independent) edges $(u_i, u_j)$ and
$(u_k, u_\ell)$ in $\mathcal{M}$ such that
$u_i < u_k < u_\ell < u_j$ or  $u_k < u_i < u_j < u_\ell$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Shuffle product on permutations ----
%%
\section{Shuffle product on permutations}
\label{section:Shuffle product on permutations}

The purpose of this section is to define a shuffle product $\SHUFFLE$
on permutations. To define this shuffle product, we shall first define
a coproduct denoted by $\Delta$, enabling to unshuffle permutations.
By duality, $\Delta$ implies the definition of $\SHUFFLE$. The reason
why we need to pass by the definition of $\Delta$ to define $\SHUFFLE$
is justified by the fact that a lot of properties of $\SHUFFLE$ depend
of properties of $\Delta$ and that this strategy allows to write concise
and clear proofs of them. We invite the reader unfamiliar with the
concepts of coproduct and duality to consult~\cite{Joni:Rota:1979}
or~\cite{Grinberg:Reiner:2014}.
\medskip

Let use denote by $\QQ[S]$ the linear span of all permutations. We
define a linear coproduct $\Delta$ on $\QQ[S]$ in the following way. For
any permutation $\pi$, we set
\begin{equation} \label{equ:unshuffling_coproduct}
    \Delta(\pi) =
    \sum_{P_1 \sqcup P_2 = [|\pi|]}
    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
\end{equation}
We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
instance,
\begin{equation}
    \Delta(213) =
    \epsilon \otimes 213 + 2. 1 \otimes 12 +
    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
\end{equation}
\begin{equation}
    \Delta(1234) =
    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
    4. 123 \otimes 1 + 1234 \otimes \epsilon,
\end{equation}
\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
    \Delta(1432) & =
    \epsilon \otimes 1432 + 3. {\bf 1 \otimes 132} + 1 \otimes 321 +
    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
    321 \otimes 1 + 1432 \otimes \epsilon.
\end{split}\end{equation}
Observe that the coefficient of the tensor $1 \otimes 132$ is $3$
in~\eqref{equ:example_unshuffling_coproduct} because there are exactly
three ways to extract from the permutation $1432$ two disjoint subwords
respectively order-isomorphic to the permutations $1$ and $132$.
\medskip

As announced, let us now use $\Delta$ to define a shuffle product on
permutations. As any coproduct, $\Delta$ leads to the definition of a
product obtained by duality in the following way.
From~\eqref{equ:unshuffling_coproduct}, for any permutation $\pi$, we
have
\begin{equation} \label{equ:schematic_coproduct}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$, denoted by
$\SHUFFLE$, is a linear binary product on $\QQ[S]$. It satisfies, for
any permutations $\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi,
\end{equation}
where the coefficients $\lambda_{\sigma, \nu}^\pi$ are the ones
of~\eqref{equ:schematic_coproduct}. We call $\SHUFFLE$ the
{\em shuffle product of permutations}. For instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3. {\bf 1432} + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
\medskip

Intuitively, this product shuffles the values and the positions of the
letters of the permutations. As first and easy properties of $\SHUFFLE$,
one can observe that the empty permutation $\epsilon$ is a unit for
$\SHUFFLE$ and that this product is graded by the sizes of the
permutations ({\em i.e.}, the product of a permutation of size $n$ with
a permutation of size $m$ produces a sum of permutations of size $n + m$.
Notice that there are other shuffle products on
permutations~\cite{DHT:IJAC:2002}: the {\em shifted shuffle} and the
{\em convolution product}. Our shuffle product is different from these
last two.
\medskip

We say that a permutation $\pi$ {\em appears} in the shuffle
$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
coefficient $\lambda_{\sigma, \nu}^\pi$ defined above is different from
zero. In a more combinatorial way, this is equivalent to say that there
are two sets $P_1$ and $P_2$ of disjoints indexes of letters of $\pi$
satisfying $P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$
is order-isomorphic to $\sigma$ and the subword $\pi_{|P_2}$ is
order-isomorphic to $\nu$.
\medskip

Besides, a permutation $\pi$ is a {\em square} if there is a
permutation $\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$.
In this case, we say that $\sigma$ is a {\em square root} of $\pi$.
Equivalently, $\pi$ is a square with $\sigma$ as square root if and only
if in the expansion of $\Delta(\pi)$, there is a tensor
$\sigma \otimes \sigma$ with a nonzero coefficient. In a more
combinatorial way, this is equivalent to say that there are two sets
$P_1$ and $P_2$ of disjoints indexes of letters of $\pi$ satisfying
$P_1 \sqcup P_2 = [|\pi|]$ such that the subword $\pi_{|P_1}$ and
$\pi_{|P_2}$ are order-isomorphic. Computer experiments give us the
first numbers of square permutations with respects to their size, which
are, from size $0$ to $10$,
\begin{equation}
    1, 0, 2, 0, 20, 0, 504, 0, 21032, 0, 1293418.
\end{equation}
For instance, the square permutations of sizes $0$ to $4$ are
\smallskip

\begin{tabular}{c|c|c}
    Size $0$ \, & Size $2$ & Size $4$ \\ \hline
    \multirow{2}{*}{\, $\epsilon$} &
    \multirow{2}{*}{\, $12$, $21$ \,} &
    $1234$, $1243$, $1423$, $1324$, $1342$, $4132$, $3124$, $3142$,
    $3412$, $4312$, \\
    & & $2134$, $2143$, $2413$, $4213$, $2314$, $2431$, $4231$, $3241$,
    $3421$, $4321$
\end{tabular}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Binary square words and permutations ----
%%
\section{Binary square words and permutations}
\label{section:Binary square words and permutations}
In this section, we shall show that the square binary words are in
one-to-one correspondence with square permutations avoiding some
patterns. This property establishes a link between the shuffle of binary
words and our shuffle of permutations and allows to obtain a new
description of square binary words.
\medskip

Let $u$ be a binary word of length $n$ with $k$ occurrences of $0$.
We denote by $\BINTOPERM$ the map sending any such word $u$ to the
permutation obtained by replacing from left to right each occurrence of
$0$ in $u$ by $1$, $2$, \dots, $k$, and from right to left each
occurrence of $1$ in $u$ by $k + 1$, $k + 2$, \dots, $n$. For instance,
\begin{equation}
    \BINTOPERM({\bf 1}00{\bf 1}0{\bf 1}{\bf 1}0{\bf 1}000) =
    {\bf C} 12 {\bf B} 3 {\bf A} {\bf 9} 4 {\bf 8} 567.
\end{equation}
Observe that for any nonempty permutation $\pi$ in the image of
$\BINTOPERM$, there is exactly one binary word $u$ such that
$\BINTOPERM(u0) = \BINTOPERM(u1) = \pi$. In support of this observation,
when $\pi$ has an even size, we denote by $\PERMTOBIN(\pi)$ the word $ua$
such that $|ua|_0$ and $|ua|_1$ are both even, where $a \in \{0, 1\}$.
\medskip

\begin{proposition} \label{prop:bijection_binary_to_permutations_squares}
    For any $n \geq 0$, the map $\BINTOPERM$ restricted to the set of
    square binary words of length $2n$ is a bijection between this last
    set and the set of square permutations of size $2n$ avoiding the
    patterns $213$ and $231$.
\end{proposition}
\begin{proof}
    [of Proposition~\ref{prop:bijection_binary_to_permutations_squares}]
    The statement of the proposition is a consequence of the
    following claims implying that $\PERMTOBIN$ is the inverse
    map of $\BINTOPERM$ over the set of square binary words.
    \begin{claim} \label{claim:binary_to_permutation_avoiding}
        The image of $\BINTOPERM$ is the set of all permutations
        avoiding $213$ and $231$.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:binary_to_permutation_avoiding}]
        Let us first show that the image of $\BINTOPERM$ contains only
        permutations avoiding $213$ and $231$. Let $u$ be a binary word,
        $\pi = \BINTOPERM(u)$, and $P_0$ (resp. $P_1$) be the set of the
        positions of the occurrences of $0$ (resp. $1$) in $u$. By
        definition of $\BINTOPERM$, from left to right, the subword
        $v = \pi_{|P_0}$ is increasing and the subword $w = \pi_{|P_1}$
        is decreasing, and all letters of $w$ are greater than those
        of~$v$. Now, assume that $\pi$ admits an occurrence of $213$.
        Then, since $v$ is increasing and $w$ is decreasing, there is an
        occurrence of $3$ (resp. $13$, $23$) in $v$ and a relative
        occurrence of $21$ (resp. $2$, $1$). All these three cases
        contradict the fact that all letters of $w$ are greater than
        those of $v$. A similar argument shows that $\pi$ avoids~$231$
        as well.
        \smallskip

        Finally, observe that any permutation $\pi$ avoiding $213$ and
        $231$ necessarily starts by the smallest possible letter or the
        greatest possible letter. This property is then true for the
        suffix of $\pi$ obtained by deleting its first letter,
        and so on for all of its suffixes. Thus, by
        replacing each letter $a$ of $\pi$ by $0$ (resp. $1$) if $a$ has the
        role of a smallest (resp. greatest) letter, one obtains a binary
        word $u$ such that $\BINTOPERM(u) = \pi$. Hence, all permutations
        avoiding $213$ and $231$ are in the image of $\BINTOPERM$.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_binary_to_square_permutation}
        If $u$ is a square binary word, $\BINTOPERM(u)$ is a square
        permutation.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_binary_to_square_permutation}]
        Since $u$ is a square binary word, there is a binary word $v$
        such that $u \in v \shuffle v$. Then, there are two disjoint
        sets $P$ and $Q$ of positions of letters of $u$ such that
        $u_{|P} = v = u_{|Q}$. Now, by definition of $\BINTOPERM$, the
        words $\BINTOPERM(u)_{|P}$ and $\BINTOPERM(u)_{|Q}$ have the
        same standarized $\rho$. Hence, and by definition of the shuffle
        product of permutations, $\BINTOPERM(u)$ appears in
        $\rho \SHUFFLE \rho$, showing that $\BINTOPERM(u)$ is a square
        permutation.
        \qed
    \end{proof}
    \begin{claim} \label{claim:square_permutation_to_square_binary}
        If $\pi$ is a square permutation avoiding $213$ and $231$,
        $\PERMTOBIN(\pi)$ is a square binary word.
    \end{claim}
    \begin{proof}[of Claim~\ref{claim:square_permutation_to_square_binary}]
        Let $\pi$ be a square permutation avoiding $213$ and $231$. By
        Claim~\ref{claim:binary_to_permutation_avoiding}, $\pi$ is in
        the image of $\BINTOPERM$ and hence, $u = \PERMTOBIN(\pi)$ is a
        well-defined binary word. Since $\pi$ is a square permutation,
        there are two disjoint sets $P_1$ and $P_2$ of indexes of letters
        of $\pi$ such that $\pi_{|P_1}$ and $\pi_{|P_2}$ are
        order-isomorphic. This implies, by the definitions of $\BINTOPERM$
        and $\PERMTOBIN$, that $u_{|P_1} = u_{|P_2}$, showing that $u$
        is a square binary word.
        \qed
    \end{proof}
\end{proof}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Algebraic issues ----
%%
\section{Algebraic issues}
\label{section:Algebraic issues}
The aim of this section is to establish some of properties of the
shuffle product of permutations $\SHUFFLE$. It is worth to note that, as
we will see, algebraic properties of the unshuffling coproduct $\Delta$
of permutations defined in
Section~\ref{section:Shuffle product on permutations} lead to
combinatorial properties of $\SHUFFLE$.
\medskip

\begin{proposition} \label{prop:shuffle_associative_commutative}
    The shuffle product $\SHUFFLE$ of permutations is associative and
    commutative.
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:shuffle_associative_commutative}]
    To prove the associativity of $\SHUFFLE$, it is convenient to show
    that its dual coproduct $\Delta$ is coassociative, that is
    \begin{equation}
        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
    \end{equation}
    where $I$ denote the identity map. This strategy relies on the fact
    that a product is associative if and only if its dual coproduct is
    coassociative. For any permutation $\pi$, we have
    \begin{equation} \begin{split}
    \label{equ:shuffle_associative_commutative}
        (\Delta \otimes I) \Delta(\pi) & =
        (\Delta \otimes I)
        \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & =
        \sum_{P_1 \sqcup P_2 = [|\pi|]} \;
        \sum_{Q_1 \sqcup Q_2 = [|P_1|]}
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
        \otimes
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
        \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \sqcup P_2 \sqcup P_3 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes
        \STD\left(\pi_{|P_2}\right) \otimes
        \STD\left(\pi_{|P_3}\right).
    \end{split} \end{equation}
    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
    is equal to the last member
    of~\eqref{equ:shuffle_associative_commutative}, whence the
    associativity of $\SHUFFLE$.
    \smallskip

    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
    if in the expansion of $\Delta(\pi)$ there is a tensor
    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
    same expansion the tensor $\nu \otimes \sigma$ with the same
    coefficient $\lambda$. Clearly, a product is commutative if and only
    if its dual coproduct is cocommutative. Now, from the
    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
    contributes to the coefficient of
    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
    the pair $(P_2, P_1)$ contributes to the coefficient of
    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
    is commutative.
    \qed
\end{proof}
\medskip

Proposition~\ref{prop:shuffle_associative_commutative} implies that
$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
coassociative cocommutative coalgebra, or in an equivalent way, that
$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
commutative algebra.
\medskip

\begin{proposition} \label{prop:endomorphisms}
    The three linear maps
    \begin{equation}
        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
    \end{equation}
    linearly sending a permutation $\pi$ to, respectively,
    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
    associative algebras.
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:endomorphisms}]
    To prove, for $j = 1, 2, 3$, that $\phi_j$ is a morphism of
    associative algebras, we have to prove that for all permutations
    $\sigma$ and $\nu$,
    \begin{equation}
        \phi_j(\sigma \SHUFFLE \nu) =
        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
    \end{equation}
    We shall proceed here as we have done during the proof of
    Proposition~\ref{prop:shuffle_associative_commutative} by showing
    that $\phi_j$ is a morphism of coalgebras, that is
    \begin{equation}
        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
    \end{equation}
    In the sequel, $\pi$ is a permutation.
    \smallskip

    If $P$ is a set of indexes of letters of $\pi$, we denote by
    $\widetilde{P}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now, since
    the operation $\widetilde{\,}$ defines a bijection on the set of the
    subsets of $[|\pi|]$, and since the standardization operation
    commutes with the mirror operation on words without multiple
    occurrence of a letter, we have
    \begin{equation} \begin{split}
        \Delta(\phi_1(\pi))
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\phi_1(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|P_1}\right)
        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \widetilde{\STD\left(\pi_{|P_1}\right)}
        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
    $\phi_1$ is a morphism of associative algebras.
    \smallskip

    Next, since by definition of the complementation operation on
    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
    we have $\tau_i < \tau_k$ if and only if $\bar \tau_i > \bar \tau_k$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_2(\pi))
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\phi_2(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\bar \pi_{|P_1}\right)
        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
    $\phi_2$ is a morphism of associative algebras.
    \smallskip

    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection, and since
    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_3(\pi))
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\phi_3(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|P_1}\right)
        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right)^{-1}
        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
        & = \sum_{P_1 \sqcup P_2 = [|\pi|]}
        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
    $\phi_3$ is a morphism of associative algebras.
    \qed
\end{proof}
\medskip

We now use the algebraic properties of $\SHUFFLE$ exhibited by
Proposition~\ref{prop:endomorphisms} to obtain combinatorial properties
of square permutations.
\medskip

\begin{proposition} \label{prop:square_stability}
    Let $\pi$ be a square permutation and $\rho$ be a square root of
    $\pi$. Then,
    \begin{enumerate}[label={\it (\roman*)},fullwidth]
        \item \label{item:square_stability_1}
        the permutation $\widetilde{\pi}$ is a square and
        $\widetilde{\rho}$ is one of its square roots;
        \item \label{item:square_stability_2}
        the permutation $\bar \pi$ is a square and $\bar \rho$ is one of
        its square roots;
        \item \label{item:square_stability_3}
        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$ is one of
        its square roots.
    \end{enumerate}
\end{proposition}
\begin{proof}[of Proposition~\ref{prop:square_stability}]
    All statements~\ref{item:square_stability_1},
    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
    consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
    since $\pi$ is a square permutation and $\rho$ is a square root of
    $\pi$, by definition, $\pi$ appears in the product
    $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
    for any $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
    one has
    \begin{equation}
        \phi_j(\rho \SHUFFLE \rho) =
        \phi_j(\rho) \SHUFFLE \phi_j(\rho).
    \end{equation}
    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
    appears in $\phi_j(\rho \SHUFFLE \rho)$ and appears also in
    $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This shows that
    $\phi_j(\rho)$ is a square root of $\phi_j(\pi)$ and
    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
    and~\ref{item:square_stability_3}.
    \qed
\end{proof}
\medskip

Let us make an observation about Wilf-equivalence classes of
permutations restrained on square permutations. The
well-known~\cite{Simion:Schmidt:EJC:1985} fact that there is a single
Wilf-equivalence class of permutations of size $3$ together with
Proposition~\ref{prop:square_stability} imply that $123$ and $321$ are
in the same Wilf-equivalence class of square permutations, and that
$132$, $213$, $231$, and $312$ are in the same Wilf-equivalence class of
square permutations. Computer experiments show us that there are two
Wilf-equivalence classes of square permutations of size $3$. Indeed, the
number of square permutations avoiding $123$ begins by
\begin{equation}
    1, 0, 2, 0, 12, 0, 118, 0, 1218, 0, 14272,
\end{equation}
while the number of square permutations avoiding $132$ begins by
\begin{equation}
    1, 0, 2, 0, 11, 0, 84, 0, 743, 0, 7108.
\end{equation}
\medskip

Besides, an other consequence of Proposition~\ref{prop:square_stability}
is that its makes sense to enumerate the sets of square permutations
quotiented by the operations of mirror image, complementary, and
inverse. The sequence enumerating these sets begins by
\begin{equation}
    1, 0, 1, 0, 6, 0, 81, 0, 2774, 0, 162945.
\end{equation}
\medskip

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Algorithmic issues ----
%%
\section{Algorithmic issues}
\label{section:Algorithmic issues}

This section is devoted to proving hardness of recognizing square permutations.
In the same way as happens with words, we shall use a linear graph framework
where deciding whether a permutation is a square reduces to computing some
specific matching in the associated linear graph \cite{Buss:Soltys:2014,Rizzi:Vialette:CSR:2013}.
It reads as follows.

\begin{lemma}
  \label{lemma:matching}
  Let $\pi$ be a permutation. The following statements are equivalent:
  \begin{enumerate}
    \item
    $\pi$ is a square.
    \item
    There exists a perfect containment-free mathing
    $\mathcal{M}$ of $G_\pi$ such that
    for any two disctinct edges
    $(a, b)$ and $(a', b')$ in $\mathcal{M}$
    we have $\pi(a) < \pi(a')$ if and only if $\pi(b) < \pi(b')$.
  \end{enumerate}
\end{lemma}

\begin{proof}[of Lemma~\ref{lemma:matching}]
    \qed
\end{proof}

Before proving hardness, we give an easy lemma that will prove extremely
useful for simplifying the proof of upcoming
Proposition~\ref{proposition:hardness}.

\begin{lemma}
  \label{lemma:at most one edge monotone}
  Let $\pi = \pi_1 \, \sigma_1 \, \pi_2 \, \sigma_2 \, \pi_3$
  be a permutation with $|\sigma_1| \geq 2$ and $|\sigma_2| \geq 2$,
  and $\mathcal{M}$ be a containment-free perfect matching of $G_\pi$.
  The following assertions hold:
  \begin{enumerate}
    \item
    If $\sigma_1$ is increasing, $\sigma_2$ is decreasing and
    $\sigma_1$ is above $\sigma_2$
    (see Fig.~\ref{subfig:increasing before and aboce decreasing}),
    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
    \item
    If $\sigma_1$ is decreasing, $\sigma_2$ is increasing and
    $\sigma_1$ is below $\sigma_2$
    (see Fig.~\ref{subfig:decresing before and below increasing}),
    then $\mathcal{M}$ sontains at most one $(\sigma_1, \sigma_2)$-edge.
  \end{enumerate}
\end{lemma}

\begin{proof}[of Lemma~\ref{lemma:at most one edge monotone}]
  \begin{figure}[t!]
    \centering
    \subfigure[%
      An increasing pattern before and above a decreasing pattern.
    ]{%
      \begin{tikzpicture}
        [
          scale=0.45,
          label/.style={anchor=base}
        ]
        % increasing pattern
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
        \foreach \x/\y in {0/6,1/7,2/8,3/9,4/10,5/11} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % decreasing pattern
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (a) at (1,5) {$a$};
        \node [label] (b) at (4,5) {$b$};
        \node [label] (ap) at (7,-1) {$a'$};
        \node [label] (bp) at (10,-1) {$b'$};
        \node (nu1) at (2.5,4) {$\sigma_1$};
        \node (nu2) at (8.5,-2) {$\sigma_2$};
        % crossings edges
        \draw [line width=1pt]
        (1,7) .. controls +(0,12) and +(0,10) .. (7,4);
        \draw [line width=1pt]
        (4,10) .. controls +(0,12) and +(0,10) .. (10,1);
      \end{tikzpicture}
      \label{subfig:increasing before and aboce decreasing}
    }% end subfigure
    \qquad
    \subfigure[%
      A decreasing pattern before and below an increasing pattern.
    ]{%
    \begin{tikzpicture}
      [
        scale=0.45,
        label/.style={anchor=base}
      ]
      % decreasing pattern
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,-0.2) grid (5.2,5.2);
      \foreach \x/\y in {0/5,1/4,2/3,3/2,4/1,5/0} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % decreasing pattern
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,5.8) grid (11.2,11.2);
      \foreach \x/\y in {6/6,7/7,8/8,9/9,10/10,11/11} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % labels
      \node [label] (a) at (1,-1) {$a$};
      \node [label] (b) at (4,-1) {$b$};
      \node [label] (ap) at (7,5) {$a'$};
      \node [label] (bp) at (10,5) {$b'$};
      \node (nu1) at (2.5,-2) {$\sigma_1$};
      \node (nu2) at (8.5,4) {$\sigma_2$};
      % crossings edges
      \draw [line width=1pt]
      (1,4) .. controls +(0,10) and +(0,12) .. (7,7);
      \draw [line width=1pt]
      (4,1) .. controls +(0,10) and +(0,12) .. (10,10);
      \end{tikzpicture}
      \label{subfig:decresing before and below increasing}
    }% end subfigure
    \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
      Lemma~\ref{lemma:at most one edge monotone}:
      Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
    }
  \end{figure}
  We first prove \emph{1}.
  Suppose, aiming at a contradiction, that there exist two
  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
  Therefore, we have
  $\pi(a) < \pi(b)$ (since $\sigma_1$ is increasing) and
  $\pi(a') > \pi(b')$ (since $\sigma_2$ is decreasing),
  which contradicts our hypothesis about $\mathcal{M}$.

  As for \emph{2}, suppose, still aiming at a contradiction,
  that there exist two
  $(\sigma_1, \sigma_2)$-edges $(a, a')$ and $(b, b')$ in $\mathcal{M}$.
  Since $\mathcal{M}$ is containment-free, we may assume $a < b$ and $a' < b'$.
  Therefore, we have
  $\pi(a) > \pi(b)$ (since $\sigma_1$ is decreasing) and
  $\pi(a') < \pi(b')$ (since $\sigma_2$ is increasing),
  which contradicts our hypothesis about $\mathcal{M}$.
  \qed
\end{proof}

\begin{proposition}
  \label{proposition:hardness}
  Deciding whether a permutation is the union of two
  order-isomorphic patterns is \NPC.
\end{proposition}

\begin{proof}
  The problem is certainly in $\NPclass$.
  We propose a reduction from the pattern involvement problem which is
  known to be \NPC~\cite{Bose:Buss:Lubiw:1998}:
  Given permutation $\pi$ and $\sigma$, decide whether $\sigma$
  occurs in $\pi$ (as an order-isomorphic pattern).

  For the sake of clarity, we will that a bunch of positions $P$
  is \emph{above} (resp. \emph{below}) above another bunch of positions $P'$
  in permutation $\mu$
  if $\mu(x) > \mu(y)$ (resp. $\mu(x) < \mu(y)$) for every $x \in P$
  and every $y \in P'$.
  For exemple, $\nu_1$ is above $\nu_2$ but below $\nu'_1$ in permutation $\mu$.

  \begin{figure}[t!]
    \centering
    \begin{tikzpicture}[
      scale=.95,
      >=stealth',
      shorten >=1pt,
      main node/.style={align=center},
      cell/.style={draw,ultra thick,fill=black!5},
      structure link/.style={line width=1.5pt},
      pattern link/.style={line width=1.5pt,black!20},
      monotone/.style={->,thick}
      ]
      % edges
      % pi 1 - pi 2
      \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
      % pi 2 - sigma 2
      \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
      % 1
      \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
      \node (claim edge 1) at (1.5,13.5) {Claim~\ref{claim:(nu_1, nu'_1)-edge}};
      \node (claim edge 1) at (1.5,13) {Claim~\ref{claim:all (nu_1, nu'_1)-edges}};
      % 2
      \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
      \node (claim edge 2) at (4,4) {Claim\ref{claim:(nu_2, nu'_2)-edge}};
      \node (claim edge 2) at (4,3.5) {Claim\ref{claim:all (nu_2, nu'_2)-edges}};
      % 3
      \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
      \node (claim edge 1) at (5.5,11.5) {Claim\ref{claim:(nu_3, nu'_3)-edge}};
      % 4
      \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
      \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
      % sigma 1 - pi 1
      \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
      \node (claim edge 1) at (6.5,4) {Claim\ref{}};
      % nodes
      % 1
      \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
      \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
      %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
      \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
      \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
      % 2
      \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
      \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
      \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
      \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
      % 3
      \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
      \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
      \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
      \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
      % 4
      \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
      \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
      \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
      \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
      % sigma 1
      \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
      \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
      % pi 1
      \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
      \node [main node] (pi1) at (8.5,5.5) {$\pi'$};
      % pi 2
      \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
      \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
      % sigma 2
      \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
      \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};
      % grid
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
      % row size
      \foreach \y/\N in {0.5/N_2,1.5/N_2,2.5/N_4,3.5/k,4.5/n,5.5/n+2,6.5/k+2,7.5/N_4,8.5/N_3,9.5/N_3,10.5/N_1,11.4/N_1} {
          \node at (-0.5,\y) (R\y) {$\N$};
      }
      % column size
      \foreach \x/\N in
      {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
          \node at (\x,-0.5) (C\x) {$\N$};
      }
    \end{tikzpicture}
    \caption{\label{fig:reduction}}
  \end{figure}

  Let $\pi \in S_n$ and $\sigma \in S_k$ be two arbitrary permutations.
  Define
  \begin{align*}
  N_4 &= 2(2n + k + 2) + 1  = 4n + 2k + 5 \\
  N_3 &= 2(2N_4 + 2n + 2k + 4) + 1 = 20n + 12k + 9 \\
  N_2 &= 2(2N_3 + 2N_4 + 2n + 2k + 4) + 1 \\
  N_1 &= 2(2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4) + 1\text{.}
  \end{align*}
  Notice that $N_1, N_2, N_3$ and $N_4$ are polynomial in $n$.
  The crucial properties are that
  (i) $N_1, N_2, N_3$ and $N_4$ are odd integers
  and
  (ii) $N_i > \left(\sum_{i < j \leq 4} 2N_j\right) + 2n + 2k + k$
  for every $1 \leq i \leq 4$.

  We now turn to defining various gadgets (sequences of integers)
  that act as building blocks in our construction of a new permutation $\mu$:
  \begin{align*}
  \sigma'  &= ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2] \\
  \pi'     &= ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2] \\
  \sigma'' &= \sigma \; [2N_2 + N_4] \\
  \pi''    &= \pi \; [2N_2 + N_4 + k] \\
  \nu_1    &= \nearrow_{N_1} \; [2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
  \nu'_1   &= \nearrow_{N_1} \; [N_1 + 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4] \\
  \nu_2    &= \nearrow_{N_2} \; [N_2] \\
  \nu'_2   &= \searrow_{N_2} \\
  \nu_3    &= \nearrow_{N_3} \; [2N_2 + 2N_4 + 2n + 2k + 4] \\
  \nu'_3   &= \nearrow_{N_3} \; [2N_2 + N_3 + 2N_4 + 2n + 2k + 4] \\
  \nu_4    &= \searrow_{N_4} \; [2N_2 + N_4 + 2n + 2k + 4] \\
  \nu'_4   &= \searrow_{N_4} \; [2N_2]\text{.}
  \end{align*}
  We are now in position to define our target permutation $\mu$:
  $$
  \mu
  =
  \nu_1 \; \nu_2 \; \nu'_1 \; \nu_3 \; \sigma' \; \nu_4 \; \nu'_3 \; \pi' \; \nu'_4 \; \pi'' \; \sigma''
  \text{.}
  $$

  It is immediate that H can be constructed in polynomial-time in $n$ and $k$.
  We claim that $\sigma$ occurs as a pattern in $\pi$ if and only if
  there exists a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.

  Suppose first that $\sigma$ occurs in $\pi$ as a pattern.
  We show how to construct a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  \begin{itemize}
    \item $\mathcal{M}$ contains $N_1$ pairwise crossing
    $(\nu_1, \nu'_1)$-edges.
    \item $\mathcal{M}$ contains $N_2$ pairwise crossing
    $(\nu_2, \nu'_2)$-edges.
    \item $\mathcal{M}$ contains $N_3$ pairwise crossing
    $(\nu_3, \nu'_3)$-edges.
    \item $\mathcal{M}$ contains $N_4$ pairwise crossing
    $(\nu_4, \nu'_4)$-edges.
    \item $\mathcal{M}$ contains $k+2$ pairwise crossing
    $(\sigma', \pi')$-edges as depicted in Figure~\ref{}.
    (Notice that we use here the fact that $\sigma$ occurs as a pattern in $\pi$).
    \item $\mathcal{M}$ contains $n-k$ pairwise crossing
    $(\pi', \pi'')$-edges as depicted in Figure~\ref{}.
    \item $\mathcal{M}$ contains $k$ pairwise crossing
    $(\pi'', \sigma'')$-edges as depicted in Figure~\ref{}.
    (Notice that we use here again the fact that $\sigma$ occurs as a pattern in $\pi$).
  \end{itemize}
  It can be easily checked (probably referring to Figure~\ref{}) that
  $\mathcal{M}$ is perfect and that
  for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.


    \begin{figure}[t!]
      \centering
        \begin{tikzpicture}
          [
            scale=0.45,
            label/.style={anchor=base},
            cell/.style={draw,thick,fill=black!5}
          ]
          % sigma'
          \draw [cell,rounded corners] (1.5,16.5) -- (4.5,16.5) -- (4.5,19.5) -- (1.5,19.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (0.8,16.8) grid (5.2,21.2);
          \foreach \x/\y in {1/20,2/19,3/17,4/18,5/21} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (pip) at (3,22) {$\sigma'$};
          \node[%
            draw,
            text width=5.5cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,-1.5cm)}
          ] at (3,26)
          {%
            $((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$
          };%
          %\node at (3,-0.5) {$((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$};
          \node at (-0.5,20) {$k+1$};
          \node at (-0.5,21) {$k+2$};
          % pi'
          \draw [cell,rounded corners] (6.5,8.5) -- (12.5,8.5) -- (12.5,14.5) -- (6.5,14.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (5.8,8.8) grid (13.2,16.2);
          \foreach \x/\y in {6/15,7/12,8/13,9/10,10/9,11/11,12/14,13/16} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (pip) at (9.5,17) {$\pi'$};
          \node[%
            draw,
            text width=5.5cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,3cm)}
          ] at (9.5,0)
          {%
            $((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$
          };%
          %\node at (9.5,-2) {$((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$};
          \node at (-0.5,15) {$n+1$};
          \node at (-0.5,16) {$n+2$};
          % pi''
          \draw [cell,rounded corners] (13.5,2.5) -- (19.5,2.5) -- (19.5,8.5) -- (13.5,8.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (13.8,2.8) grid (19.2,8.2);
          \foreach \x/\y in {14/6,15/7,16/4,17/3,18/5,19/8} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (pipp) at (16.75,9) {$\pi''$};
          \node[%
            draw,
            text width=2.25cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,-4cm)}
          ] at (16.5,19)
          {%
            $\pi [2N_2 + N_4 + k]$
          };%
          %\node at (16.5,-3.5) {$\pi [2N_2 + N_4 + k]$};
          % sigma''
          \draw [cell,rounded corners] (19.5,-0.5) -- (22.5,-.5) -- (22.5,2.5) -- (19.5,2.5) -- cycle;
          \draw[step=1cm,black!60,ultra thin,fill=black!10] (19.8,-0.2) grid (22.2,2.2);
          \foreach \x/\y in {20/2,21/0,22/1} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          \node (sigmapp) at (21,3) {$\sigma''$};
          \node[%
            draw,
            text width=1.75cm,text justified,color=black!50,rounded corners,
            rectangle callout,callout relative pointer={(0.0cm,1.25cm)}
          ] at (21,-4.5)
          {%
            $\sigma [2N_2 + N_4]$
          };%
          %\node at (21,-5) {$\sigma [2N_2 + N_4]$};
          % edges sigma' - pi'
          \draw [line width=1pt]
          (1,20) .. controls +(0,9) and +(0,4) .. (6,15);
          \draw [line width=1pt]
          (5,21) .. controls +(0,9) and +(0,4) .. (13,16);
          \draw [line width=1pt]
          (2,19) .. controls +(0,-17) and +(0,-7) .. (8,13);
          \draw [line width=1pt]
          (3,17) .. controls +(0,-17) and +(0,-7) .. (9,10);
          \draw [line width=1pt]
          (4,18) .. controls +(0,-17) and +(0,-7) .. (11,11);
          % edges pi' - pi''
          \draw [line width=1pt]
          (7,12) .. controls +(0,9) and +(0,4) .. (14,6);
          \draw [line width=1pt]
          (10,9) .. controls +(0,9) and +(0,4) .. (17,3);
          \draw [line width=1pt]
          (12,14) .. controls +(0,9) and +(0,4) .. (19,8);
          % edges pi'' - sigma''
          \draw [line width=1pt]
          (15,7) .. controls +(0,-9) and +(0,-4) .. (20,2);
          \draw [line width=1pt]
          (16,4) .. controls +(0,-9) and +(0,-4) .. (21,0);
          \draw [line width=1pt]
          (18,5) .. controls +(0,-9) and +(0,-4) .. (22,1);
          % text
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(0.5cm,-.75cm)}
          % ] at (1,27)
          % {%
          % $2$ pairwise crossing $(\sigma',\pi')$-edges that
          % link the guards
          % };%
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(-0.1cm,-2.5cm)}
          % ] at (17,24)
          % {%
          % $n-k$ pairwise crossing $(\sigma',\pi')$-edges that
          % denote positions that do not correspond to
          % a specific occurence of $\sigma$ in $\pi$
          % (at positions $2$, $3$ and $5$).
          % };%
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(0.75cm,1.5cm)}
          % ] at (2.5,4,5)
          % {%
          % $k$ pairwise crossing $(\sigma',\pi')$-edges that
          % denote a specific occurence of $\sigma$ in $\pi$
          % (at positions $2$, $3$ and $5$).
          % };%
          % \node[%
          %   draw,
          %   text width=5.5cm,text justified,color=black!50,
          %   rectangle callout,callout relative pointer={(2cm,.75cm)}
          % ] at (7,-3)
          % {%
          % $k$ pairwise crossing $(\pi''; \sigma'')$-edges that
          % denote a specific occurence of $\sigma$ in $\pi$
          % (at positions $2$, $3$ and $5$).
          % };%
        \end{tikzpicture}
        \caption{\label{fig:subfig:sigma' - pi' - pi'' - sigma''}%
        $\sigma = 312$ and $\pi = 4\mathbf{5}\mathbf{2}1\mathbf{3}6$
        (where a specific occurrence of $\sigma$ in $\pi$ is depicted in bold).
        }
      \end{figure}

  More precisely,
  (i) the first integer of $\sigma'$
  (\emph{i.e.}, $(k+1) + (2N_2 + N_4 + 2n + k + 2)$) is linked
  to the first integer of $\pi'$
  ((\emph{i.e.}, $(n+1) + (2N_2 + N_4 + n + k + 2)$)),
  (ii)

  Conversely, suppose that there exists a containment-free perfect matching
  $\mathcal{M}$ in $G_\mu$ such that for any two disctinct edges
  $(a, b)$ and $(a', b')$ in $\mathcal{M}$
  we have $\mu(a) < \mu(a')$ if and only if $\mu(b) < \mu(b')$.
  We show that $\sigma$ occurs as a pattern in $\pi$.
  Whereas the matching $\mathcal{M}$ may not be as regular as in the forward
  direction, the main idea is to prove that $\mathcal{M}$ contains enough structure
  (more precisely, enough $(\sigma', \pi')$-edges) so that we can conclude.
  We have divided the reverse direction into a serie of basic claims that
  progressively defines the overall structure of $\mathcal{M}$.

    \begin{figure}[t!]
      \centering
      \subfigure[%
        A $(\nu_1, \nu_2)$-edge $(a, b)$ together with a
        $(\nu_2, \nu'_1)$-edge $(x, y)$.
        Notice that we may have $x<b$
        (as long as $N_1 < x \leq N_1+N_2$ and $N_1 < b \leq N_1+N_2$
        do hold).
      ]{%
        \begin{tikzpicture}
          [
            scale=0.25,
            label/.style={anchor=base}
          ]
          % nu_1
          \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
          \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          % nu_4
          \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
          \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          % nu'_1
          \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
          \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
              \draw [fill=black] (\x,\y) circle (0.1);
          }
          % labels
          \node [label] (y) at (17,14.5) {$y$};
          \node [label] (yp) at (10,-1.5) {$y'$};
          \node [label] (a) at (0,8.5) {$a$};
          \node [label] (ap) at (8,-1.5) {$a'$};
          \node (nu1) at (2.5,7.7) {$\nu_1$};
          \node (nu2) at (8.5,-2.5) {$\nu_2$};
          \node (nu1') at (14.5,13.5) {$\nu'_1$};
          % crossings edges
          \draw [line width=1pt]
          (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
          \draw [line width=1pt]
          (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
          \draw [line width=1pt]
          (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
          \draw [line width=1pt]
          (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
          \draw [line width=1pt]
          (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
          % data edges
          \draw [line width=1pt]
          (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
          \draw [line width=1pt]
          (10,1) .. controls +(0,32) and +(0,10) .. (17,21);
        \end{tikzpicture}
        \label{subfig:no (nu_1, nu_2)-edge - 1}
      }% end subfigure
      \qquad
      \subfigure[%
        A $(\nu_1, \nu'_1)$-edge $(x', y')$ together with an
        edge $(y, z)$ which is neither
        a $(\nu_1, \nu'_1)$-edge,
        nor a $(\nu_2, \nu'_1)$-edge,
        nor a $(\nu'_1, \nu'_1)$-edge.
        Notice that $y$ may be any position satisfying
        $N_1+N_2 < y \leq 2N_1+N_2$.
      ]{%
      \begin{tikzpicture}
        [
          scale=0.25,
          label/.style={anchor=base}
        ]
        % nu_1
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
        \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu_4
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
        \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu'_1
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
        \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (y) at (17,14.5) {$y$};
        \node [label] (ypp) at (23,-4.5) {$y''$};
        \node [label] (a) at (0,8.5) {$a$};
        \node [label] (a) at (2,8.5) {$x'$};
        \node [label] (a) at (14,14.5) {$y'$};
        \node [label] (ap) at (8,-1.5) {$a'$};
        \node (nu1) at (2.5,7.7) {$\nu_1$};
        \node (nu2) at (8.5,-2.5) {$\nu_2$};
        \node (nu1') at (14.5,13.5) {$\nu'_1$};
        % crossings edges
        \draw [line width=1pt]
        (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
        \draw [line width=1pt]
        (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
        \draw [line width=1pt]
        (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
        \draw [line width=1pt]
        (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
        \draw [line width=1pt]
        (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
        % data edges
        \draw [line width=1pt]
        (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
        \draw [line width=1pt]
        (17,21) .. controls +(0,10) and +(0,35) .. (23,-3);
        \end{tikzpicture}
        \label{subfig:no (nu_1, nu_2)-edge - 2}
      }% end subfigure
      \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
        Claim~\ref{claim:no (nu_1, nu_2)-edge}:
        Assuming a $(\nu_1,\nu_2)$-edge $(a, b)$ in $\mathcal{M}$.
      }
    \end{figure}

  \begin{claim}
    \label{claim:no (nu_1, nu_2)-edge}
    There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_1, \nu_2)$-edge $e = (a, a')$ in $\mathcal{M}$.
    Under this assumption, we show that $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Indeed, suppose, aiming at a contradiction,
    that there exists a $(\nu_1, \nu_1)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free,
    we must have $x < a$.
    Then it follows that $\mu(x) < \mu(a)$ (since $\nu_1$ is increasing) and
    $\mu(x') > \mu(a')$ (since $\nu_1$ is above $\nu_2$),
    which contradicts our hypothesis about $\mathcal{M}$.
    Hence $\mathcal{M}$ does not contain any $(\nu_1, \nu_1)$-edge.
    Furthermore, since $\nu_1$ is increasing and
    there exists only one $(\nu_1, \nu_2)$-edge $(a, a')$ in $\mathcal{M}$,
    we conclude that
    $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
    But $|\nu_1| = |\nu'_1| = N_1$ and $\mathcal{M}$ is perfect,
    and hence there exists a position, say $y$, in $\nu'_1$ that is
    not involved in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    We rule out this configuration by considering two cases:
    \begin{itemize}
      \item Position $y$ is involved in a $(\nu_2, \nu'_1)$-edge,
      say $(y', y)$. Then, we have $\mu(a) > \mu(y')$ (since $\nu_1$ is above $\nu_2$)
      and $\mu(a') < \mu(y)$ (since $\nu_2$ is below $\nu'_1$),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 1}.
      \item Position $y$ is not involved in a $(\nu_2, \nu'_1)$-edge.
      Since $\mathcal{M}$ contains $N_1-1$ pairwise crossing $(\nu_1, \nu'_1)$-edges
      in $\mathcal{M}$ and $|\nu'_1| = N_1$, there exists an edge $(y, y'')$ in $\mathcal{M}$
      that is neither a $(\nu_1, \nu'_1)$-edge nor a $(\nu_2, \nu'_1)$-edge
      nor a $(\nu'_1, \nu'_1)$-edge.
      In other words, $y'' > 2N_1 + N_2$.
      Now, for any $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$, we
      have $\mu(x) < \mu(y)$ (since $\nu_1$ is below $\nu'_1$)
      and $\mu(x') > \mu(y'')$ (since $\nu'_1$ is above any other gadget),
      which contradicts our hypothesis about $\mathcal{M}$.
      See Fig.~\ref{subfig:no (nu_1, nu_2)-edge - 2}.
    \end{itemize}
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_1, nu'_1)-edge}
    There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_1, nu'_1)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Then it follows that there exists an edge $(x, x')$ in $\mathcal{M}$,
    $1 \leq x \leq N_1$, that is neither a
    $(\nu_1, \nu_1)$-edge (since $N_1$ is odd)
    nor a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_1, \nu'_1)$ (by our contradiction hypothesis).
    (In other words, $x \leq N_1$ and $x' > 2N_1 + N_2$.)
    Therefore, since $\mathcal{M}$ is containment-free,
    there is neither a $(\nu_2, \nu_2)$-edge nor a $(\nu'_1, \nu'_1)$-edge
    in $\mathcal{M}$.
    Then it follows that every position $y$ of $\nu'_1$
    (\emph{i.e.}, $N_1+N_2 < y \leq 2N_1+N_2$)
    is involved in an edge $(y, y'$ of $\mathcal{M}$ with $y' \geq 2N_1+N_2$
    (otherwise $\mathcal{M}$ would not be containment-free).
    But $|\nu'_1| = N_1 = 2(2N_2 + 2n_3 + 2N_4 + 2n + 2k + 4) + 1
    > 2N_2 + 2N_3 + 2N_4 + 2n + 2k + 4$, and hence at least one
    position in $\nu'_1$ can not be involved in an edge in $\mathcal{M}$.
    Then it follows that $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  The above claim will be complement in upcoming Claim~\ref{claim: all (nu_1, nu'_1)-edges}.

  \begin{claim}
    \label{claim:no (nu_2, nu_2)-edge}
    There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_2)-edge}]
  Combine Claim~\ref{claim:(nu_1, nu'_1)-edge} with
  the fact that $\mathcal{M}$ is containment-free.
  \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_2, nu'_1)-edge}
    There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu'_1)-edge}]
    First, according to Lemma~\ref{lemma:at most one edge monotone},
    there exists at most one $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
    Suppose now, aiming at a contradiction, that there exists
    exactly one $(\nu_2, \nu'_1)$-edge $e = (a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_1, nu'_1)-edge}, there exists at least
    one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
    Furtheremore, since
    there is no $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge}),
    every position in $\nu_1$ is involved either in a
    $(\nu_1, \nu_1)$-edge or in a $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$
    (otherwise $\mathcal{M}$ would not be containment-free).
    Then it follows that that there is an odd number of $(\nu_1, \nu'_1)$-edges
    in $\mathcal{M}$ since $\mathcal{M}$ is perfect and $|\nu_1| = N_1$ is odd.
    Combining this with the fact that there exists exactly one $(\nu_2, \nu'_1)$-edge
    in $\mathcal{M}$ and $|\nu'_1| = N_1$ is odd,
    we conclude that there exists a position, say $x$, in $\nu'_1$
    that is neither involved in a $(\nu_1, \nu'_1)$-edge nor a $(\nu_1, \nu'_1)$-edge
    nor a $(\nu'_1, \nu'_1)$-edge.
    Write $(x, x')$, $N_1+N_2 < x \leq 2N_1 + N_2$ and $x' > 2N_1 + N_2$, this edge.
    We have $\mu(a) < \mu(x)$ (since $\nu_2$ is below $\nu'_1$)
    and $\mu(a') > \mu(x')$ (since $\nu'_1$ is above any other gadget),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_2, nu'_2)-edge}
    There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_2, nu'_2)-edge}]
    Suppose, aiming at a contradiction that there is no
    $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
    Notice that there is neither
    a $(\nu_1, \nu_2)$-edge (Claim~\ref{claim:no (nu_1, nu_2)-edge})
    nor a $(\nu_2, \nu_2)$-edge (Claim~\ref{claim:no (nu_2, nu_2)-edge})
    nor a $(\nu_2, \nu'_1)$-edge (Claim~\ref{claim:no (nu_2, nu'_1)-edge})
    in $\mathcal{M}$.
    But $|\nu_2| = N_2 > 2N_3 + 2N_4 + 2n + 2k +4$.
    Hence $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no edge below (nu_2, nu'_2)-edge}
    There is
    neither a $(\nu'_1, \nu'_1)$-edge,
    nor a $(\nu'_1, \nu_3)$-edge,
    nor a $(\nu'_1, \sigma')$-edge,
    nor a $(\nu'_1, \nu_4)$-edge
    nor a $(\nu_3, \nu_3)$-edge,
    nor a $(\nu_3, \sigma')$-edge,
    nor a $(\nu_3, \nu_4)$-edge,
    nor a $(\sigma', \sigma')$-edge,
    nor a $(\sigma', \nu_4)$-edge,
    nor a $(\nu_4, \nu_4)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
    Combine Claim~\ref{claim:(nu_2, nu'_2)-edge} with the fact that
    $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim: no edge right of nu'_2}
    There is
    neither a $(\nu'_2, \nu'_3)$-edge,
    nor a $(\nu'_2, \pi')$-edge,
    nor a $(\nu'_2, \nu'_4)$-edge,
    nor a $(\nu'_2, \pi'')$-edge
    nor a $(\nu'_2, \sigma'')$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim: no edge right of nu'_2}]
    Suppose, aiming at a contradiction, that
    $\mathcal{M}$ contains an edge $(a, a')$ which is
    either a $(\nu'_2, \nu'_3)$-edge
    or a $(\nu'_2, \pi')$-edge
    or a $(\nu'_2, \nu'_4)$-edge
    or a $(\nu'_2, \pi'')$-edge
    or a $(\nu'_2, \sigma'')$-edge.
    According to Claim~\ref{claim:(nu_2, nu'_2)-edge},
    there exists a $(\nu_2, \nu'_2)$-edge $(b, b')$ in $\mathcal{M}$.
    Then it follows that we have
    $\mu(b) > \mu(a)$ (since $\nu_2$ is above $\nu'_2$) and
    $\mu(b') < \mu(a')$ (since $\nu'_3$, $\pi'$, $\nu'_4$, $\pi''$ and
    $\sigma''$ are all above $\nu'_2$),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  % \begin{claim}
  %   \label{claim:no (nu_1, nu'_2)-edge}
  %   There is no $(\nu_1, \nu'_2)$-edge in $\mathcal{M}$.
  % \end{claim}
  %
  % \begin{proof}[of Claim~\ref{claim:no (nu_1, nu'_2)-edge}]
  %   Suppose, aiming at a contradiction that there is a
  %   $(\nu_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
  %   Since $\mathcal{M}$ is containment-free, we must have $b < a$.
  %   Then it follows that we have
  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is increasing) and
  %   $\mu(b') > \mu(a')$ (since $\nu'_2$ is above $\nu'_2$),
  %   which contradicts our hypothesis about $\mathcal{M}$.
  %   \qed
  % \end{proof}
  %
  % \begin{claim}
  %   \label{claim:no (nu'_1, nu'_2)-edge}
  %   There is no $(\nu'_1, \nu'_2)$-edge in $\mathcal{M}$.
  % \end{claim}
  %
  % \begin{proof}[of Claim~\ref{claim:no (nu'_1, nu'_2)-edge}]
  %   Suppose, aiming at a contradiction that there is a
  %   $(\nu'_1, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
  %   According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
  %   there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
  %   Then it follows that we have
  %   $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
  %   $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu'_2$),
  %   which contradicts our hypothesis about $\mathcal{M}$.
  %   \qed
  % \end{proof}

  \begin{figure}[t!]
    \centering
      \begin{tikzpicture}
        [
          scale=0.35,
          label/.style={anchor=base}
        ]
        % nu_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,5.8) grid (5.2,11.2);
        \foreach \x/\y in {0/11,1/10,2/9,3/8,4/7,5/6} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu_3
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,15.8) grid (11.2,21.2);
        \foreach \x/\y in {6/16,7/17,8/18,9/19,10/20,11/21} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % nu'_2
        \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,-0.2) grid (17.2,5.2);

        \foreach \x/\y in {12/5,13/4,14/3,15/2,16/1,17/0} {
            \draw [fill=black] (\x,\y) circle (0.1);
        }
        % labels
        \node [label] (x) at (2,4.5) {$x$};
        \node [label] (a) at (8,14.5) {$a$};
        \node [label] (xp) at (14,-1.5) {$x'$};
        \node [label] (ap) at (16,-1.5) {$a'$};
        \node (nu1) at (2.5,2.5) {$\nu_2$};
        \node (nu2) at (8.5,12.5) {$\nu_3$};
        \node (nu1') at (14.5,-3.5) {$\nu'_2$};
        % crossings edges
        \draw [line width=1pt]
        (2,9) .. controls +(0,25) and +(0,30) .. (14,3);
        \draw [line width=1pt]
        (8,18) .. controls +(0,15) and +(0,20) .. (16,1);
      \end{tikzpicture}
    \caption{\label{fig:subfig:no (nu_3, nu'_2)-edge}%
      Claim~\ref{claim:no (nu_3, nu'_2)-edge}:
      Assuming a $(\nu_3, \nu'_2)$-edge $(a, b)$ in $\mathcal{M}$.
    }
  \end{figure}

  \begin{claim}
    \label{claim:no (nu_3, nu'_2)-edge}
    There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_3, nu'_2)-edge}]
    Suppose, aiming at a contradiction, that there exists
    $(\nu_3, \nu'_2)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_2, nu'_2)-edge}, there
    exists at least one $(\nu_2, \nu'_2)$-edge $(x, x')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free (Proposition~\ref{proposition:matching}),
    we have $x' < a'$.
    Therefore,
    we have $\mu(x) < \mu(a)$ (since $\nu_2$ is below $\nu_3$) and
    $\mu(x') > \mu(a')$ (since $\nu'_2$ is decreasing),
    which contradicts our hypothesis about $\mathcal{M}$.
    See Fig.~\ref{fig:subfig:no (nu_3, nu'_2)-edge}.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:at most one (nu_2, nu_3)-edge}
    There is at most one $(\nu_2, \nu_3)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:at most one (nu_2, nu_3)-edge}]
    Apply Lemma~\ref{lemma:at most one edge monotone}.
    \qed
  \end{proof}

    We will see soon (upcoming Claim~\ref{claim:all (nu_2, nu'_2)-edges})
    that there exists actually no $(\nu_2, \nu_3)$-edge
    in $\mathcal{M}$.

  \begin{claim}
    \label{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}
    There is neither a $(\nu_1, \nu_3)$-edge not a $(\nu'_1, \nu_3)$-edge
    in $\mathcal{M}$.
  \end{claim}

 \begin{proof}[of Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge}]
    We first prove that there is no $(\nu_1, \nu_3)$-edge in $\mathcal{M}$.
    Indeed, suppose, aiming at a contradiction, that these exists
    a $(\nu_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free, we must have
    $b < a'$.
    Then it follows that
    $\mu(b) < \mu(a)$ (since $\nu_1$ is incresing) and
    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
    which contradicts our hypothesis about $\mathcal{M}$.

    As for no $(\nu_1, \nu_3)$-edges in $\mathcal{M}$, suppose,
    aiming at a contradiction, that these exists
    a $(\nu'_1, \nu_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_1, nu'_1)-edge},
    there exists a $(\nu_1, \nu'_1)$-edge $(b, b')$ in $\mathcal{M}$.
    Then it follows that
    $\mu(b) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
    $\mu(b') > \mu(a')$ (since $\nu'_1$ is above $\nu_3$),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_3, nu'_3)-edge}
    There exists a $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_3, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_3, \nu'_3)$-edge in $\mathcal{M}$.
    Combining
    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
    Claim~\ref{claim:no (nu_3, nu'_2)-edge},
    Claim~\ref{claim:at most one (nu_2, nu_3)-edge}
    and Claim~\ref{claim:no (nu_1, nu_3)-edge and no (nu'_1, nu_3)-edge},
    we conclude that $N_3-1$ positions in $\nu_3$ are involved
    in edges of $\mathcal{M}$ that are
    neither ($\nu_1, \nu_3)$-edges,
    nor $(\nu_2, \nu_3)$-edges
    nor $(\nu'_1, \nu_3)$-edges
    nor $(\nu_3, \nu_3)$-edges.
    But $N_3 - 1 > 2N_4 + 2n + 2k + 4$ (the positions left) and hence
    $\mathcal{M}$ is not a perfect matching,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no edge below one (nu_3, nu'_3)-edge}
    There is
    neither a $(\sigma', \sigma')$-edge,
    nor a $(\sigma', \nu_4)$-edge,
    nor a $(\sigma', \nu'_2)$-edge,
    nor a $(\nu_4, \nu_4)$-edge
    nor a $(\nu_4, \nu'_2)$-edge,
    nor a $(\nu'_2, \nu'_2)$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge}]
    Combine Claim~\ref{claim:(nu_3, nu'_3)-edge} with the fact that
    $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  The next two claims state that $\mathcal{M}$ actually contains
  all $(\nu_1, \nu'_1)$-edges and all $(\nu_2, \nu'_2)$-edges.

  \begin{claim}
    \label{claim:all (nu_1, nu'_1)-edges}
    $\mathcal{M}$ contains $N_1$ pairwise crossing $(\nu_1, \nu'_1)$-edges.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:all (nu_1, nu'_1)-edges}]
    Suppose, aiming at a contradiction, that
    $\mathcal{M}$ does not contain $N_1$ $(\nu_1, \nu'_1)$-edge.
    Combing Claim~\ref{claim:no (nu_2, nu'_1)-edge} and
    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge},
    we conclude that there exists a position $a$ in $\nu'_1$ that is
    involded in an edge $(a, a')$ in $\mathcal{M}$ that is
    neither a $(\nu_1, \nu'_1)$-edge,
    nor a $(\nu_2, \nu'_1)$-edge,
    nor a $(\nu'_1, \nu'_1)$-edge.
    (In other words, $a' > 2N_1 + N_2$.)
    Furthermore, according to Claim~\ref{claim:(nu_1, nu'_1)-edge},
    there exists a $(\nu_1, \nu'_1)$-edge $(x, x')$ in $\mathcal{M}$.
    Then, we have
    $\mu(x) < \mu(a)$ (since $\nu_1$ is below $\nu'_1$) and
    $\mu(x') < \mu(a')$ (since $\nu'_1$ is above any other gadget),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:all (nu_2, nu'_2)-edges}
    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:all (nu_2, nu'_2)-edges}]
    The key idea is to focus on $\nu'_2$.
    Combining Claim~\ref{claim:no (nu_3, nu'_2)-edge},
    Claim~\ref{claim: no edge right of nu'_2} and
    Claim~\ref{claim:no edge below one (nu_3, nu'_3)-edge},
    we obtain that every position in $\nu'_2$ is involved in
    either a $(\nu_1, \nu'_2$)-edge
    or a $(\nu'_1, \nu'_2$)-edge
    or $(\nu_2, \nu'_2$)-edge.
    But, according to Claim~\ref{claim:all (nu_1, nu'_1)-edges},
    all positions in $\nu_1$ and $\nu'_1$ are involved in
    $(\nu_1, \nu'_1)$-edges.
    Then it follows that
    all positions in $\nu'_2$ are involved in a $(\nu_2, \nu'_2)$,
    and hence
    $\mathcal{M}$ contains $N_2$ pairwise crossing $(\nu_2, \nu'_2)$-edges.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_2, nu_4)-edge}
    There is no $(\nu_2, \nu_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_2, nu_4)-edge}]
    According to Claim~\ref{claim: all (nu_2, nu'_2)-edges},
    all positions in $\nu_2$ and $\nu'_2$ are involved in
    $(\nu_2, \nu'_2)$-edges in $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (nu_4, nu'_3)-edge}
    There is no $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no (nu_4, nu'_3)-edge}]
    Suppose, aiming at a contradiction, that there exists
    a $(\nu_4, \nu'_3)$-edge $(a, a')$ in $\mathcal{M}$.
    According to Claim~\ref{claim:(nu_3, nu'_3)-edge},
    there exists a $(\nu_3, \nu'_3)$-edge $(x, x')$
    in $\mathcal{M}$.
    Since $\mathcal{M}$ is containment-free, we must have
    $x' < a'$.
    Therefore,
    we have $\mu(x) > \mu(a)$ (since $\nu_3$ is above $\nu_4$) and
    $\mu(x') < \nu(a')$ (since $\nu'_3$ is increasing),
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:(nu_4, nu'_4)-edge}
    There is at least one $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:(nu_4, nu'_4)-edge}]
    Suppose, aiming at a contradiction, that there is no
    $(\nu_4, \nu'_4)$-edge in $\mathcal{M}$.
    First, according to
    Claim~\ref{claim:all (nu_1, nu'_1)-edges} and
    Claim~\ref{claim:all (nu_2, nu'_2)-edges},
    there is neither a $(\nu_1, \nu_4)$-edge
    nor a $(\nu'_1, \nu_4)$-edge
    nor a $(\nu_2, \nu_4)$-edge
    nor a $(\nu_4, \nu'_2)$-edge in $\mathcal{M}$.
    Furthermore, according to
    Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge} and
    Claim~\ref{claim:no (nu_4, nu'_3)-edge},
    there is neither a $(\nu_3, \nu_4)$-edge
    nor a $(\nu_4, \nu'_3)$-edge in $\mathcal{M}$.
    But $N_4 > 2n + 2k + 2$, and
    hence $\mathcal{M}$ is not perfect,
    which contradicts our hypothesis about $\mathcal{M}$.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no edge below (nu_4, nu'_4)-edge}
    There is
    neither a $(\pi', \pi')$-edge
    nor a $(\sigma', \pi'')$-edge
    nor a $(\sigma', \sigma'')$-edge
    in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\ref{claim:no edge below (nu_2, nu'_2)-edge}]
    Combine Claim~\ref{claim:(nu_4, nu'_4)-edge} with the
    fact that $\mathcal{M}$ is containment-free.
    \qed
  \end{proof}

  \begin{claim}
    \label{claim:no (sigma', nu'_3)-edge, no (sigma', nu'_4)-edge}
    There is neither a $(\sigma', \nu'_3)$-edge
    nor a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
  \end{claim}

  \begin{proof}[of Claim~\label{claim:no (sigma', nu'_3)-edge}]
    According to
    Claim~\ref{claim: no edge right of nu'_2},
    Claim~\ref{claim:all (nu_1, nu'_1)-edges}
    Claim~\ref{claim:all (nu_2, nu'_2)-edges} and
    Claim~\ref{claim:no (nu_4, nu'_3)-edge}
    every position in $\sigma'$ is involved in
    either a $(\sigma', \nu'_3)$-edge
    or a $(\sigma', \pi')$-edge
    or a $(\sigma', \nu'_4)$-edge
    in $\mathcal{M}$.
    Let $a$ be the first position of $\sigma'$
    (\emph{i.e.}, $a = 2N_1 + N_2 + N_3 + 1$)
    and
    $b$ be the last position of $\sigma'$
    (\emph{i.e.}, $b = 2N_1 + N_2 + N_3 + k + 2$),
    and write $(a, a')$ and $(b, b')$ for the two associated edges of $\mathcal{M}$.
    We consider the presence of a $(\sigma', \nu'_3)$-edge
    or a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$ separately.
    \begin{itemize}
      \item
      Suppose, aiming at a contradiction, that there exists
      a $(\sigma', \nu'_3)$-edge in $\mathcal{M}$.
      Since $\mathcal{M}$ is containment-free, we conclude that
      $(a, a')$ must be a $(\sigma', \nu'_3)$-edge.
      But $\mu(a) = 2N_2 + N_4 + 2n + k + 2 + (k+1) <
      \mu(b) = 2N_2 + N_4 + 2n + k + 2 + (k+2)$, and hence
      $(b, b')$ is also a $(\sigma', \nu'_3)$-edge
      (otherwise we would have $\mu(a') > \mu(b')$ since
      $\nu'_3$ is above $\pi'$ and $\nu'_4$,
      which contradicts our hypothesis).
      Consider any position $c$ but $a$ and $b$ of $\sigma'$
      (\emph{i.e.},
      $a = 2N_1 + N_2 + N_3 + 1 < c < 2N_1 + N_2 + N_3 + k + 2 = b$),
      and let $(c, c')$ be the associate edge in $\mathcal{M}$.
      We first observe that $(c, c')$ is not a $(\sigma', \nu'_3)$-edge,
      since otherwise we would have
      $\mu(a) > \mu(c)$ and $\mu(a') < \mu(c')$
      (since $\mathcal{M}$ is containment-free and $\nu'_3$ is increasing),
      thereby contradicting our hypothesis about $\mathcal{M}$.
      Therefore, $(c, c')$ is either a $(\sigma', \pi)$-edge or a $(\sigma', \nu'_4)$-edge.
      In both cases, $(b, b')$ is contained in $(c, c')$,
      which contradicts our hypothesis about $\mathcal{M}$ being containment-free.
      \item
      Suppose, aiming at a contradiction, that there exists
      a $(\sigma', \nu'_4)$-edge in $\mathcal{M}$.
      Since $\mathcal{M}$ is containment-free, we conclude that
      $(b, b')$ must be a $(\sigma', \nu'_4)$-edge.
      Since there is no $(\sigma', \nu'_3)$-edge in $\mathcal{M}$
      (preceding point),
      $(a, a')$ is either a
      $(\sigma', \pi')$-edge or a $(\sigma', \nu'_4)$-edge.
      In both cases, we have $a' < a$.
      Then it follows that
      $\mu(a) < \mu(b)$
      (since $2N_2 + N_4 + 2n + k + 2 + (k+1) < 2N_2 + N_4 + 2n + k + 2 + (k+2)$)
      and
      $\mu(a') > \mu(b')$ (since $\pi'$ is above $\nu'_4$ if $(a, a')$ is
      a $(\sigma', \pi')$-edge and
      $\nu'_4$ is decreasing if $(a, a')$ is
      a $(\sigma', \nu'_4)$-edge),
      which contradicts our hypothesis about $\mathcal{M}$.
    \end{itemize}
    \qed
  \end{proof}

  Combining the above claims, we conclude that there are
  $k+2$ $(\sigma', \pi')$-edges in $\mathcal{M}$.
  Recall that
  $\sigma' = ((k+1) \; \sigma \; (k+2)) \; [2N_2 + N_4 + 2n + k + 2]$ and that
  $\pi' = ((n+1) \; \pi \; (n+2)) \; [2N_2 + N_4 + n + k + 2]$.
  Then it follows we have
  at least $k$ (possibly $k+1$ or $k+2$) independent $(\sigma', \pi')$-edges
  $(a, a')$ in $\mathcal{M}$ with
  $2N_1 + N_2 + N_3 + 1 < a < 2N_1 + N_2 + N_3 + k + 2$
  and
  $2N_1 + N_2 + N_3 + (k + 2) + 1 < a' < 2N_1 + N_2 + N_3 + (k + 2) + n + 2$.
  Therefore, by our hypothesis about $\mathcal{M}$,
  $\sigma$ occurs as a pattern in $\pi$.
  \qed
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% %%
% %% ---- Patterns ----
% %%
% \section{Patterns}
% \label{section:Patterns}
%
% \begin{proposition}
%   \label{proposition:patterns:lower bound}
%   Let $\pi \in S_{n}$.
%   Then $\pi$ contains a pattern of size at least
%   $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
%   is the union of two order-isomorphic patterns.
% \end{proposition}
%
% \begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
%   The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935}
%   states that every permutation of
%   length at least $pq+1$ must contain either the pattern
%   $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
%   The result now follows from the fact that
%   the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
% \qed
% \end{proof}
%
% \begin{proposition}
%   \label{proposition:patterns:upper bound}
%   If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
%   a permutation of $S_{n}$
%   that does not contains a pattern
%   of length $2k$ that is the union of two
%   order-isomorphic patterns.
% \end{proposition}
%
% \begin{proof}[of Proposition~\ref{proposition:patterns:upper bound}]
% The proof is by the probabilistic method \cite{Alon:Spencer:1992}.
% Let $\pi \in S_n$ be chosen randomly and uniformly.
% For any pattern $\sigma$ of length $2k$ ($k$ to be precisely defined latter)
% of $\pi$,
% let $X_\sigma$ be the indicator variable that
% $\sigma$ is the union of two order-ismorphic patterns.
% According to Boole's inequality, we have
% $P\left(X_\sigma\right) \leq \binom{2k}{k}\;2^{-k}$.
% Since there are $\binom{n}{2k}$ patterns of length $2k$ in $\pi$,
% the probability that $\pi$ contains a pattern of length $2k$
% is the union of two order-ismorphic patterns is at most
% $\binom{n}{2k}\;\binom{2k}{k}\;2^{-k} < 1$.
% Thus, with positive probability, no event $X_\sigma$ occurs and
% there is a permutation of $S_n$ that does not contain a pattern
% of length $2k$ that is the union of two order-isomorphic patterns.
% \qed
% \end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% ---- Conclusion ----
%%
\section{Conclusion}
\label{section:Conclusion}

There are a number of further directions of investigation in this
general subject. We mention several -~not necessarily new~- open
problems that are, in our opinion, the most interesting. How many
permutations of $S_{2n}$ are squares? How many $(213,231)$-avoiding
permutations of $S_{2n}$ are squares? (Equivalently, how many binary
strings of length $2n$ are squares; see Problem~4
in \cite{Henshall:Rampersad:Shallit:2011})? How hard is the problem of
deciding whether a $(213,231)$-avoiding permutation is a square
(see Problem~4 in \cite{Henshall:Rampersad:Shallit:2011})?) Given two
permutations $\pi$ and $\sigma$, how hard is the problem of deciding
whether $\pi$ is the union of two patterns order-isomorphic to $\sigma$?
\smallskip

Besides, in this work, we investigate some algebraic properties of the
shuffle of permutations which have direct combinatorial consequences.
However, this work raises several additional classical algebraic
questions. Among these, one can ask for a complete algebraic study of
$\QQ[S]$ as a graded associative algebra for the shuffle  product
$\SHUFFLE$. Describing a generating family for $\QQ[S]$, defining
multiplicative bases of $\QQ[S]$, and determining whether $\QQ[S]$ is
free as an associative algebra are worthwhile questions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Bibliography
%%

\bibliographystyle{plain}
\bibliography{biblio}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section*{Appendix (Reviewers' version only)}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
